A large-scale study of polymorphism in R

### 2019-03-17 paper/main.tex
working on the results section

\begin{abstract}
The R programming language is widely used in a variety of scientfic domains
for tasks related to data science. The language was designed to favor an
interactive style of programming with minimal syntactic and conceptual
overhead. This design is well suited to support interactive data analysis,
but is not well suited to generating performant code or catching programming
errors.  In particular, R has no type annotations and all operations are
dynamically checked at runtime. The starting point for our work is the
question: \emph{what could a static type system for R look like?}  To answer
that question we study the polymorphism that is present in over X millions
of lines of R code, XX packages, written over a period of Y years by Z
programmers.  We perform a dynamic analysis, leveraging tests and use-cases,
to determine the level of polymorphism that is present in the code. We do
this for several potential notions of types. Our result suggest that
polymorphism is important in some key parts of the system but that
relatively simple type annotations could be used to capture most of the
interesting cases.
\end{abstract}
\maketitle

@@

Our community builds, improves, and reasons about programming languages.  To
make design decisions that benefit most users of a language, we need to
understand the language we are working with as well as the real-world needs
it answers. Often, we, as researchers, can appeal to our intuition as many
languages are intended to be general purpose and appeal to users with some 
training.

@@

To faciliate programming these language offer all kinds of short cuts and
facilities that increase the likelyhood that code will ``just work''.

@@

The results obtained are published and even sometimes drive policy decisions.

@@

In spite of these flaws, the language is incredibly popular.
\AT{We'll want to cite something here.}

@@

@@

@@

\AT{Tables, graphs, code examples should go in here.}
In this section, we ... .

>>>>

\begin{abstract}
The R programming language is widely used in a variety of scientific domains
for tasks related to data science. The language was designed to favor an
interactive style of programming with minimal syntactic and conceptual
overhead. This design is well suited to support interactive data analysis,
but is not well suited to generating performant code or catching programming
errors.  In particular, R has no type annotations and all operations are
dynamically checked at runtime. The starting point for our work is the
question: \emph{what could a static type system for R look like?}  To answer
that question we study the polymorphism that is present in over X millions
of lines of R code spread among some 7500 packages, written over a period of over 20 years by over 10000
programmers.  We perform a dynamic analysis, leveraging tests and use-cases,
to determine the level of polymorphism that is present in the code. We do
this for several potential notions of types. Our result suggest that
polymorphism is important in some key parts of the system but that
relatively simple type annotations could be used to capture most of the
interesting cases.
\end{abstract}

\maketitle

@@

Our community builds, improves, and reasons about programming languages.  To
make design decisions that benefit most users of a language, we need to
understand the language we are working with as well as the real-world needs
it answers. Often, we, as researchers, can appeal to our intuition as many
languages are intended to be general purpose and appeal to users with some 
training. Unfortunately, these intuitions don't always apply to domain-specific
languages.

@@

To facilitate programming these language offer all kinds of short cuts and
facilities that increase the likelihood that code will ``just work''.

@@

The results obtained are published and even sometimes drive policy decisions.
\AT{There's a cool stat in the blog I linked which says that a lot of the Stack
Overflow traffic R gets is from government officials.}

@@

In spite of these flaws, the language is incredibly popular.
\AT{We'll want to cite something here, maybe the blog I linked?}

@@

The main contributions of this work are:

	\item \AT{maybe} a tool to analyze the polymorphicity of R code;

	\item a study of polymorphism in the R language \AT{using said tool};

	\item a number of proposed static type systems for R which cover most of the 
	type usage patterns which arise in R code.

@@

There are a number of interesting packages in our corpus.
\AT{References for all of these.}
For example, we have analyzed some packages for plotting data: among others, the {\tt ggplot2} package is an immensely popular package for generating highly-customizable plots, and the {\tt plotmo} package is an interesting package aimed at displaying models.
We also analyzed {\tt dplyr}, an integral package to the {\tt tidyverse} set of R packages \AT{talk about tidyverse earlier, or here}.
{\tt dplyr} is a popular package that makes full use of R's features to create what is tantamount to a small DSL for data manipulation within R.
Another interesting package is {\tt rquery}, a package for generating SQL-like queries.

@@

\AT{PSA:} There are a few ways we could do this section.
One way would be to go through our data in a sort of linear fashion, akin to how the main notebook is set up.
We can explore each of our type systems one at a time, and present them (thus breaking the section up by type system).
Another option is to just break down by-result, so have a section for the counts, for the top polymorphism...
I like the first option better.

\AT{Draft}
In this section, we will present the results of our analysis. 
We approach the data with the purpose of informing the design of a static type system for R.
We propose to begin with a very basic type system, discussed next.

\subsection{Base R Type System}

The {\it Base R Type System} (BRTS) ...
